Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming

### 2017-04-21 docs/reviews/review_response.tex

% headers
\usepackage{fancyhdr}
\pagestyle{fancy}
\chead{} 
\rhead{\thepage} 
% footer
\lfoot{\small\scshape } 
\cfoot{} 
%%%% insert your name here %%%%
% \rfoot{\footnotesize Author} 
\renewcommand{\headrulewidth}{.3pt} 
\renewcommand{\footrulewidth}{.3pt}
\setlength\voffset{-0.25in}
\setlength\textheight{648pt}

\begin{document}

\title{Response to referees \\
C034405: "Solar-Electric and Gas Powered, Long-Endurance UAV Sizing via Geometric Programming``}
\author{Warren Hoburg and Michael Burton}
\maketitle

Note to the editor:

The peer review of this manuscript was insightful and helpful. At a high level, the reviewers had comments about the clarity of our assumptions, objective, and approach.  We have added material that explicitly discusses the requirements against which we are comparing the two aircraft, charts explaining our approach and process, and configuration drawings and text to explain some of our assumptions.  Reviewers also stated that the manuscript was long.  We have cut unnecessary figures and equations to help shorten it.  Overall feedback was positive with minor additional questions that we have responded to herein.  Only one referee asked that this manuscript be submitted as an engineering note.  Given the overall positive feedback and our responses to the reviewers comments we respectfully submit that this manuscript be reconsidered as full paper instead of an engineering note.

@@

We thank the anonymous referees for their thoughtful and insightful comments on the draft manuscript.
We feel that the paper has improved significantly thanks to the feedback. 
Specific comments from the reviews (in italics) are addressed herein.

\section*{Referee 1}

\emph{
This article presents a description of a methodology to design a hale type aircraft using geometric programming.
As such, the approach is probably of more relevance than the specifics.
It lacks focus. Frankly this article is far more appropriate as a conference paper than a journal article. It reads like a recipe with a significant number of assumptions that are not supported.
While the approach is definitely of value, the presentation is less so. If published, this article should be shortened to an engineering note. }

We thank the referee for this comment. We have made organizational changes to the paper, including a more detailed description of the requirements, GP approach, and results. As with any aircraft design, we made many assumptions and explicitly stated them.  We argue that a great benefit to the GP approach is the ability to change those assumptions and to resolve the problem quickly.  We do this in the results section of the paper and believe this shows important insights about long-endurance UAVs. We would be happy to respond to any sepecific assumptions that the referee feels are not supported.

\emph{
A flow chart/figure could be included that shows the parameters (equations) that feed into the formulation. This way the reader gets the “bigger” picture. }

We thank the referee for this helpful comment.  We have since added a flow chart.

\emph{
The most significant figures to include are the results only. }

We thank the referee for this comment.  We did cut multiple figures.

\emph{
Perhaps show a sketch of what a typical vehicle would look like. }

We thank the referee for this comment.  We included a vehicle sketch.

\emph{
Include only the most relevant equations.
}

We thank the referee for this comment.  We did cut a few unnecessary equations to help shorten the paper.

### 2017-04-21 docs/paperin_review.tex
change fitting style; more edits; better flowchart

\begin{longtable}{lccccccccccccc}
\caption{Mission Requirements}\\
\toprule
\toprule
\label{t:mreqs}
Payload & 10 lbs\\
Endurance & 6 days\\
Season & all seasons\\
Altitude & > 4,600 m\\
Latitude & $\pm30^{\circ}$\\
\bottomrule
\end{longtable}

\subsection{Payload}

A 10 lbs payload is a baseline requirement for both the gas and solar-electric powered aircraft. This requirement can be changed to observe how each architeure is able to handle larger or smaller payloads.

Inherent in a payload requirement is the ability to station keep. To station keep an aircraft must fly at least as fast as the wind speed,

@@

The wind speed at a station is a function of latitude, altitude, season or day of the year, and percentile wind speed,

@@

For example, an aircraft optimized for $\pm$30 degrees latitude will have 11 different wind speed and solar flux constraints that must be satisfied, one for each degree of latitude between 20 and 30 degrees. 
Because wind speeds are low and solar flux is higher below 20 degrees latitude, no constraints for latitudes below 20 degrees are actually included in the optimization. 
The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the year 2005-2015.
In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees. (See Appendix A) 
The constraint on December wind speeds and air density at the 30 degree latitude is

\begin{align}
    \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
    \label{e:windfit35}
    &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
\end{align}

where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed. The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
    \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
	\label{f:windfitl35}
	\end{center}
\end{figure}

@@

where $\eta_{\text{motor}} = 0.95$. Thus for a given latitude and day, and therfore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude. 
% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
% 
% \begin{figure}[H]
%  \begin{subfigmatrix}{2}% number of columns
%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
%  \end{subfigmatrix}
%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
%  \label{f:energyapprox}
% \end{figure}

@@

\subsection{Steady Level Flight}

@@

\subsection{Aerodynamics}

@@

\subsection{Wing Spar Model}

@@

\subsubsection{Standard Wing Bending Case}

@@

\subsubsection{Gust Loading Case}

@@

\subsubsection{Discretized Beam Model}

@@

\subsubsection{Cap Spar for Bending Loads}

@@

\subsection{Wing Skin Weight}

@@

\subsection{Empennage}

@@

The drag of the horizontal and vertical tails is computed using a GP-compatible fit of XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,

\begin{align}
    D_{\text{(v,h)}} &\geq \frac{1}{2} c_{d_{\text{(v,h)}}} \rho V^2 S_{\text{(v,h)}} \\
    \label{e:taildrag}
    c_{d_{\text{(v,h)}}}^{70.5599} &\geq \num{7.42688d-90} \left( \frac{Re_{\text{(v,h)}}}{\num{1d3}} \right)^{-33.0637}(100 \tau_{\text{(v,h)}})^{18.0419}  \nonumber \\
                 & + \num{5.02826d-163}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-18.7959} (100\tau_{\text{(v,h)}})^{53.1879} \\
                 &+ \num{4.22901d-77}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-41.1704} (100\tau_{\text{(v,h)}})^{28.4609} \nonumber \\
\end{align}

where the selected airfoil is the NACA 0008 for both the horizontal and vertical tails (i.e. $\tau_{\text{(v,h)}} = 0.08$). 
The XFOIL data was generated for a zero angle of attack, based upon steady level flight where neither surface is generating lift.  
% A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.

@@

>>>>

\begin{longtable}{lccccccccccccc}
\caption{Mission Requirements}\\
\toprule
\toprule
\label{t:mreqs}
Payload & 10 lbs\\
Station Keeping & 90\% \\
Endurance & $>6$ days\\
Season & all seasons\\
Altitude & $>$ 4,600 m\\
Latitude & $\pm30^{\circ}$\\
\bottomrule
\end{longtable}

\subsection{Station Keeping}

To station keep an aircraft must fly at least as fast as the wind speed,

@@

Observations in wind speed data\cite{wind} indicate that it is unlikely for long-endurance UAVs to be able to station keep in 100 percentile wind speeds.  
Therefore, the station keeping requirement is defined as a percentage which is the percentile wind speed. 
The wind speed at a station is a function of latitude, altitude, season or day of the year, and percentile wind speed,

@@

For example, an aircraft optimized for $\pm$30 degrees latitude will have 11 different wind speed and solar flux constraints that must be satisfied, one for each degree of latitude between 20 and 30 degrees. 
Because wind speeds are low and solar flux is higher below 20 degrees latitude, no constraints for latitudes below 20 degrees are actually included in the optimization. 
The wind speed data used to generate these equations includes December wind speeds from the Northern Hemisphere and June wind speeds from the Southern Hemisphere for the year 2005-2015.
In total, 41 wind speed constraints were generated to represent $\pm$20-60 degrees each with a log space RMS error of less than 5\%. (See Appendix A) 
% The constraint on December wind speeds and air density at the 30 degree latitude is 
% 
% \begin{align}
%     \left(\frac{V_{\text{wind}}}{V_{\text{ref}}}\right)^{7.066} &\geq 18413.2\rho^{9.3}p_{\text{wind}}^{5.51} + 9658.5\rho^{8.25}p_{\text{wind}}^{23.47} \nonumber \\
%     \label{e:windfit35}
%     &+ \num{2.55d-13}\rho^{-7.72}p_{\text{wind}}^{49.65} + \num{1.11d-16}\rho^{-9.80}p_{\text{wind}}^{10.26} + 10.493.8\rho^{7.31}p_{\text{wind}}^{226.9}
% \end{align}
% 
% where $V_{\text{ref}} = 100$ [m/s] and $p_{\text{wind}}$ is the percentile wind speed. The comparison of the fitted Equation~\eqref{e:windfit35} to the actual data is presented in Figure~\ref{f:windfitl35}. 
% 
% \begin{figure}[H]
% 	\begin{center}
% 	\includegraphics[width=0.6\textwidth,natwidth=511,natheight=405]{windfitl30.pdf}
%     \caption{\textbf{Posynomial approximation to wind speed data\cite{wind} (circles) at 35$^{\circ}$ latitude. (Log-space RMS error = 0.043.)}}
% 	\label{f:windfitl35}
% 	\end{center}
% \end{figure}

@@

where $\eta_{\text{motor}} = 0.95$. Thus for a given latitude and day, and therfore total solar energy $(E/S)_{\text{sun}}$, both $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ are functions of $(P/S)_{\text{min}}$.  
The total solar energy available $(E/S)_{\text{sun}}$, is computed prior to the optimization solve and is used to generate monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$ as functions $(P/S)_{\text{min}}$, one for each degree of latitude between 20 and 60 degrees latitude. 
$(E/S)_{\text{sun}}$, $(E/S)_{\text{day}}$, $(E/S)_{\text{twilight}}$, and $t_{\text{night}}$ are precomputed using the formulas in Appendix B to generate the fitted equations. 
% Figure~\ref{f:energyapprox} shows sample monomial approximations at 30 degrees latitude. 
% 
% \begin{figure}[H]
%  \begin{subfigmatrix}{2}% number of columns
%      \subfigure[Monomial Approxmiation for $(E/S)_{\text{day}}$\label{f:Benergy}]{\includegraphics[natwidth=576,natheight=432]{Benergy.pdf}}
%      \subfigure[Monomial Approximation for $(E/S)_{\text{twilight}}$\label{f:Cenergy}]{\includegraphics[natwidth=575,natheight=432]{Cenergy.pdf}}
%  \end{subfigmatrix}
%  \caption{\textbf{Monomial approximations for $(E/S)_{\text{day}}$ and $(E/S)_{\text{twilight}}$. (Log-space RMS error = 0.007 and 0.0180)}}
%  \label{f:energyapprox}
% \end{figure}

@@

\subsubsection{Steady Level Flight}

@@

\subsubsection{Aerodynamics}

@@

\subsubsection{Wing Spar Model}

@@

\begin{itemize}

\item[Standard Wing Bending Case]

@@

\item[Gust Loading Case]

@@

\item[Discretized Beam Model]

@@

\item[Cap Spar for Bending Loads]

@@

\end{itemize}

\subsubsection{Wing Skin Weight}

@@

\subsubsection{Empennage}

@@

The drag of the horizontal and vertical tails is computed using a GP-compatible fit of XFOIL data for a range of Reynolds numbers and NACA airfoil thicknesses,
where the selected airfoil is the NACA 0008 for both the horizontal and vertical tails (i.e. $\tau_{\text{(v,h)}} = 0.08$). 
The XFOIL data was generated for a zero angle of attack, based upon steady level flight where neither surface is generating lift.

% \begin{align}
%     D_{\text{(v,h)}} &\geq \frac{1}{2} c_{d_{\text{(v,h)}}} \rho V^2 S_{\text{(v,h)}} \\
%     \label{e:taildrag}
%     c_{d_{\text{(v,h)}}}^{70.5599} &\geq \num{7.42688d-90} \left( \frac{Re_{\text{(v,h)}}}{\num{1d3}} \right)^{-33.0637}(100 \tau_{\text{(v,h)}})^{18.0419}  \nonumber \\
%                  & + \num{5.02826d-163}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-18.7959} (100\tau_{\text{(v,h)}})^{53.1879} \\
%                  &+ \num{4.22901d-77}\left(\frac{Re_{\text{(v,h)}}}{\num{1d3}}\right)^{-41.1704} (100\tau_{\text{(v,h)}})^{28.4609} \nonumber \\
% \end{align}

% A comparison of the XFOIL data and Equation~\eqref{e:taildrag} is shown in Figure~\ref{f:taildragpolar}.

@@

\section*{Appendix D}

\subsection{Solar energy calculations}
